---
title: "STA 305 HW54"
author: "Jacky Guan"
date: "2023-11-24"
output: pdf_document
---


Loading Required libraries
```{r setup, message=FALSE}
library(tidyverse)
library(openintro)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

```{r setwd, message=FALSE, eval = FALSE}
setwd("Documents/SEM\ 3/STA\ 305/Homework/HW5")
```

## 1. Regular expressions







### a) Use the words data set, find all the words that match the following pattern:


* are exactly four letters long

```{r four}
words_4 <- words %>%
  tibble() %>%
  filter(nchar(.) == 4)

print(words_4)
```



* are either four or five letters long


```{r four/five}
words_45 <- words %>%
  tibble() %>%
  filter(nchar(.) == 4 | nchar(.) == 5)

print(words_45)
```

* the second letter is “s” or “t”

```{r s/t}
s_t <- words %>%
  tibble() %>%
  filter(str_detect(words, "^.{1}[st]"))

print(s_t)
```



* contains the pattern like “oxx” where “o” is one letter and “x” is another letter 

```{r oxx}
oxx <- words %>%
  tibble() %>%
  filter(str_detect(words, "o(.)\\1"))

print(oxx)
```


* contains “a”, “e” and “o” at the same time

```{r aeo}
aeo <- words %>%
  tibble() %>%
  filter(str_detect(words, "^(?=.*a)(?=.*e)(?=.*o).*"))

print(aeo)
```




### b) Use the sentences data set, make the following plot


* a bar plot counting sentences with and without “the” (or “The”).

```{r the}
the <- sentences %>%
  str_detect("(the|The)") %>%
  tibble() %>%
  rename(Contains_The = ".")

ggplot(the) +
  geom_bar(aes(Contains_The, fill = Contains_The)) +
  theme_minimal() +
  labs(x = "Contains the/The in the Sentence",
       y = "Number of Sentences",
       title = "Sentences Containing \"The\"",
       fill = "Legend") +
  geom_text(
    aes(x = factor(Contains_The), label = after_stat(count)),
    stat = "count",
    vjust = -0.5,
    color = "black",
    size = 4
  )
```


* a scatterplot with x being the average length of words in a sentence, and y being the number of words starting with “a” or “e” or “i” or “o” or “u” in the sentence.

```{r scatterplot}
sentence_summary <- sentences %>%
  tibble() %>%
  rename(sentence = ".") %>%
  mutate(word_count = str_count(sentence, " ") + 1,
         vowel_start = str_count(sentence, "\\b[aeiouAEIOU]")) %>%
  ggplot() +
  geom_point(aes(word_count, vowel_start), position = "jitter") +
  theme_minimal() +
  labs(x = "Word Count",
       y = "Number of Words beginning with a Vowel",
       title = "Vowel as the First Letter vs Number of Words in Sentence")

sentence_summary
```



### c) Application


#### i) Download the Oxford English Dictionary as a “.txt” file from https://canvas.feitian.edu/files/9699/ download?download_frd=1

Done

#### ii) Read it into RStudio with read_lines() function (check how to use it by yourself)

```{r read_dict}
oxford <- tibble(read.delim("Oxford_English_Dictionary.txt"))
```

#### iii) Turn the dictionary into a tibble and remove all blank lines


#### iv) Use regular expression to extract all words for each item in a separate column named “words”


```{r first_word}
oxford <- oxford %>%
  rename(definition = A) %>%
  mutate(word = str_extract(definition, "\\b\\w+")) %>%
  select(word, definition)
```


## 2. Factors

```{r read_bank}
bank <- tibble(read.csv("BankChurners.csv"))
```

### a) Use the BankChurners.csv to answer the following questions:

* Which features can be regarded as a factor?

Factor features include Attrition_Flag, Gender, Dependent_count, Education_Level, 
Marital_Status, Income_Category, and Card_Category.

* Which features can be regarded as an ordered factor (ordinal)?

Of the aforementioned factors, Education_Level, Income_Category, Dependent_count, 
and Card_Category can be considered ordered factors.

* Read BankChurners.csv into RStudio, then change the columns that you answered above into factors or ordered factors.

```{r change_names}
bank <- bank %>%
  mutate(across(c(Attrition_Flag, Gender, Marital_Status), as.factor)) %>%
  mutate(Dependent_count = factor(Dependent_count, ordered = TRUE,
                                  levels = c("0", "1", "2", "3",
                                             "4", "5", "6"))) %>%
  mutate(Education_Level = factor(Education_Level, ordered = TRUE,
                                  levels = c("Uneducated", "High School",
                                             "College", "Graduate",
                                             "Post-Graduate", "Doctorate",
                                             "Unknown"))) %>%
  mutate(Income_Category = factor(Income_Category, ordered = TRUE,
                                  levels = c("Less than $40K", "$40K - $60K",
                                             "$60K - $80K", "$80K - $120K",
                                             "$120K +"))) %>%
  mutate(Card_Category = factor(Card_Category, ordered = TRUE,
                                levels = c("Blue", "Silver", "Gold")))
```



* Visualize the effect of education level on Average Utilization Ratio

```{r edu_ratio, fig.width=10}
ggplot(bank) +
  facet_grid(~ Education_Level) +
  geom_histogram(aes(Avg_Utilization_Ratio, fill = Education_Level)) +
  theme_linedraw() +
  labs(x = "Utilization Ratio",
       y = "Count",
       title = "Effect of Education Level on Average Utilization Ratio",
       fill = "Education Level") +
  theme(axis.text.x = element_blank())
```




### b) Use the gss_cat data set


* What are the levels of marital variable?


The levels of marital status include the following:

```{r gss_cat}
levels(gss_cat$marital)
```



* Combine “Separated”, “Divorced”, “Widowed” into a new category “Once Married” 

```{r once_married}
gss_cat <- gss_cat %>%
  mutate(marital = case_when(
    marital == "Separated" ~ "Once Married",
    marital == "Divorced" ~ "Once Married",
    marital == "Widowed" ~ "Once Married",
    TRUE ~ marital
  ))
```


* Use the new levels, explore whether there is an effect of martial status on tvhours.

It seems that there is a plausible correlation with being once married and the 
increse in time spent in front of a tv. For people once married, the median time 
spent in front of a tv seems to be greater compared to unmarried and married individuals.

```{r tv_married, fig.width=10}
ggplot(gss_cat) +
  facet_grid(~ marital) +
  geom_histogram(aes(tvhours, fill = marital)) +
  theme_linedraw() +
  labs(x = "TV Hours",
       y = "Count",
       title = "Effect of Marital Status on Hours Spent Watching TV",
       fill = "Marital Status")
```
