---
title: "STA 305 HW 3"
author: "Jacky Guan"
date: "2023-10-03"
output: pdf_document
---

Loading required libraries


```{r setup, message=FALSE}
library(tidyverse)
library(openintro)
library(ggridges)
library(nycflights13)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```


## 1. In-class Exercise for flights data set


### (a) filter - filter the data set and answer how many flights in the data set satisfy the given criteria.


#### • Find all flights that travel a distance of less than 200 miles.

 <br>   

It seems that there are only a little less than 23,000 flights in 2013 that traveled a distance of less than 200 miles. 
All of these flights didn't stay in the air for longer than 2 hours. 

```{r flights200}
flights200 <- flights %>%
  filter(distance <= 200) %>%
  select(tailnum, dest, air_time) %>%
  arrange(desc(air_time))

tibble(flights200)

nrow(flights200)
```


#### • Find all flights that depart from JFK and land on ORD or CVG (Chicago or Dallas).

There were 3,323 flights that went from JFK to ORD or CVG, meaning that about 9 flights departed per day that went to either airport. 

```{r JfkChicagoDallas}
chicago_or_dallas <- flights %>%
  filter(origin == "JFK", dest == "ORD" | dest == "CVG")

tibble (chicago_or_dallas)

nrow(chicago_or_dallas)
nrow(chicago_or_dallas) / 365
```


#### • Find all flights that flied a distance between 200 and 500 miles.

There are 62677 flights that flew between 200 and 500 miles in 2013.

```{r flights2_5}
flights200_500 <- flights %>%
  filter(between(distance, 200, 500))

tibble(flights200_500)

nrow(flights200_500)
```

#### • Find all flights that have a missing plane tail number. What is the percentage of flights with a missing plane tail number?

There are 2,512 planes missing a tail number.
This corresponds to about 0.7% of all flights. 
This might be due to some form of system outage perhaps caused by electromagnetic interference.

```{r missing_tail}
missing_tail <- flights %>%
  filter(is.na(tailnum))

tibble(missing_tail)

nrow(missing_tail)

perc_missing_tail <- nrow(missing_tail) * 100 / nrow(flights)
print(paste0(perc_missing_tail, "%"))
```

### (b) select and rename


#### • Reorder the columns of flights such that origin, destination, travel distance, and air time are placed in the first four columns. 
Rename the travel distance column to be “travel_distance” (this question was not originally included in class exercise).


```{r select_and_rename}
select_and_rename <- flights %>%
  mutate(distance, travel_distance = distance) %>%
  select(origin, dest, travel_distance, air_time, everything())

tibble(select_and_rename)
```

### (c) arrange


#### • What was the longest travel distance for any flight in our data set? What was the origin and the destination?

The longest flights were from `JFK` to `HNL`. 
```{r longest_flight}
longest_flight <- flights %>%
  arrange(desc(distance))

tibble(longest_flight)
```

What is the full expression for `HNL`? It is Honolulu International Airport.

```{r hnl}
filter(airports, faa == "HNL")
```

#### • What was the shortest air time for any flight (that actually finished the trip) in our data set? What was the origin and the destination?

The shortest flight only lasted for 20 minutes and flew from Newark International airport to `BDL`, which stands for Bradley International Airport. 
```{r shortest_air_time}
shortest_flight <- flights %>%
  filter(!is.na(air_time)) %>%
  arrange(air_time)

tibble(shortest_flight)

filter(airports, faa == "BDL")
```

### (d) mutate

#### • Find the flight with the highest average speed. What is its destination?

It looks like a flight going from LaGuardia International airport flying all the way to Hartsfield Jackson Atlanta International Airport was in quite the hurry. 
The plane had an average speed of 703 miles/hour, which is already faster than the speed of sound. I hope there is some sort of measurement mistake here. 
The more I read about it, the less normal this seems to me. 

```{r avg_speed}
average_speed <- flights %>%
  mutate(speed = distance/air_time*60) %>%
  arrange(desc(speed))

tibble(average_speed)

filter(airports, faa == "ATL")
```

#### • For diamonds data, create a new variable named carat_group, which groups data by less than 1 carat, 1-2 carat, 2-3 carat and greater than 3 carat.

It seems that most diamonds found are less than 1 carat. 

```{r diamond_carat_group}
diamond_carat_group <- diamonds %>%
  mutate(carat_category = case_when(
    between(carat, 0, 1) ~ "< 1 carat",
    between(carat, 1, 2) ~ "1-2 carat",
    between(carat, 2, 3) ~ "2-3 carat",
    between(carat, 3, 9) ~ "3+ carat",
    .default = NA
  ))
 
ggplot(diamond_carat_group) +
  geom_bar(aes(x = carat_category, fill = carat_category)) +
  labs(x = "Carat Category",
       title = "Number of diamonds in each carat category.")
```

### (e) group_by and summarize


#### • Find which airport around NYC had the worst delay on average.

The data is telling us that Newark has the worst delays averaging at around 15 minutes. JFK comes at a close second with 12 minutes on average while 
LaGuardia is doing well with only an average of 10 minutes of delay per flight.

```{r worst_delay}
worst_delay <- flights %>%
  group_by(origin) %>%
  summarize(delay = mean(dep_delay, na.rm = TRUE))

tibble(worst_delay)
```

#### • Find which airport around NYC had the worst average delay on December 24, 2013.

On Christmas Eve 2013, Newark still managed to have the worst delays. 

```{r worst_delay_christmas}
worst_delay_christmas <- flights %>%
  filter(month == 12, day == 24) %>%
  group_by(origin) %>%
  summarize(delay = mean(dep_delay, na.rm = TRUE))

tibble(worst_delay_christmas)
```

#### • Find which origin has the most destinations?

It looks like Newark International Airport has the most destinations out of the three that are in the city. 

```{r most_dest}
destinations <- flights %>%
  group_by(origin) %>%
  summarize(n_dest = dest[!duplicated(dest)])

tibble(destinations)

jfk <- sum(destinations$origin == "JFK")
ewr <- sum(destinations$origin == "EWR")
lga <- sum(destinations$origin == "LGA")

most_dest <- data.frame(
  origin = c("JFK", "EWR", "LGA"),
  airports = c(jfk, ewr, lga)
)

tibble(most_dest)
```

## 2. Data Transformation and Visualization with the flights data set


Working with the data set flights in the package nycflights13, answer the following questions by performing necessary data transformation/visualization.


### (a) Create a histogram of arrival delays (excluding NAs) for all flights in June and July. Comment on your findings.

The results show that July definitely has more delays compared with June in 2013. 
At the same time, July also seems to have more early arrivalals compared with June. 
This might be due to the amount of traffic that flows through airports 
during summer vacation months. There might just be more flights in July compared with June 
such that the emperical number of delays and early arrivals go up.

```{r arrival_delays}
arrival_delays <- flights %>%
  filter(between(month, 6, 7))

ggplot(arrival_delays) +
  geom_histogram(aes(x = arr_delay, fill = as.factor(month)), bins = 90) +
  xlim(-60, 300) +
  labs(x = "Delay in minutes", y = "Number of flights ",
       title = "Histogram of delayed flights in June and July",
       subtitle = "Seperated by color",
       fill = "Month")
```

### (b) Create a smooth line graph of arrival delays vs departure delays for all flights departing from EWR on the first day of each month. Comment on your findings.

For most months, it seems that arrival and departure delays are linearly related to each other. 
This is evident by the lines running throuhg each graph. It is no surprise that the number of 
delays increases during the months of July and August due to those months being the high seasons. 


However, it seems that as the departure delay time increases, the arrival delay time seems to 
become less. In other words, the derivative of the graph of arrival vs departure delay is a 
decreasing function. This might be explained throught that when flights depart a few hours after 
their intended departure time, they have more "buffer room" to speed up and catch up with their 
planned flight time.  


```{r first_day_delays}
first_day_delays <- flights %>%
  filter(day == 1)

ggplot(first_day_delays, aes(x = dep_delay,
  y = arr_delay, color = as.factor(month))) +
  geom_point() +
  geom_smooth() +
  xlim(-20, 300) +
  ylim(-60, 300) +
  facet_wrap(~ month) +
  labs(x = "Departure Delay in minutes", y = "Arrival Delay in minutes",
       title = "Dot plots of Arrival and Departure on the first day of each Month",
       fill = "Month")
```

### (c) Find the flights that actually departed with the shortest travel distance. What is its origin and destination airport?

From this graph, it seems that the shortest flight out of NYC by travel distance is headed towards 
Philadelphia. The flights take about 30 minutes on average and leave from both JFK and Newark

```{r shortest_distance}
shortest_distance <- flights %>%
  select(origin, dest, distance, air_time) %>%
  arrange(distance) %>%
  filter(!is.na(air_time))

shortest_distance  <- shortest_distance[1:500, ]

ggplot(shortest_distance) +
  geom_boxplot(aes(x = dest, y = air_time, fill = origin)) +
  labs(x = "Destination airport of shortest flight", y = "Air time in minutes",
       title = "Analysis of the 500 shortest flights in 2013",
       fill = "Origin")
```

### (d) Create a new categorical variable with two labels. Flights with a travel distance shorter than 500 miles are marked as “short-distance”, and otherwise “long-distance”. Create a bar plot to compare the number of flights in each category.


It seems that when comparing short- and long-distance flights, there are about 
3x more long-distance flights compared with short distance ones.

```{r short_vs_long}
short_vs_long <- flights %>%
  mutate(flight_distance_category = case_when(
    distance < 500 ~ "short-distance",
    distance >= 500 ~ "long-distance"
  ))

ggplot(short_vs_long, aes(x = flight_distance_category)) +
  geom_bar(position = "dodge", aes(fill = as.factor(month))) +
  labs(x = "Category of flight", y = "Number of flights ",
       title = "Number of short- vs long-distance flights from NYC in 2013",
       fill = "Distance category")
```

### (e) Find the destination airport that has the longest average departure delay by creating a graph.

It looks like `CAE` or Columbia Metropolitan Airport has the longest delays.


```{r longest_delay}
longest_delay <- flights %>%
  group_by(dest) %>%
  select(dest, dep_delay) %>%
  mutate(avg_delay = mean(dep_delay, na.rm = TRUE)) %>%
  distinct(dest, .keep_all = TRUE) %>%
  arrange(desc(avg_delay)) %>%
  filter(avg_delay > 20)


ggplot(longest_delay) +
  geom_boxplot(aes(dest, avg_delay))

filter(airports, faa == "CAE")
```

### (f) Answer the question in (d) without creating a graph.

It looks like `CAE` or Columbia Metropolitan Airport has the longest delays.


```{r longest_delay_1}
longest_delay_1 <- flights %>%
  group_by(dest) %>%
  select(dest, dep_delay) %>%
  mutate(avg_delay = mean(dep_delay, na.rm = TRUE)) %>%
  distinct(dest, .keep_all = TRUE) %>%
  arrange(desc(avg_delay))

tibble(longest_delay_1[1:10, ])
```

### (g) Find the carriers with the highest and the lowest average flight speed for all their flights in the data set.

The fastest carrier is `HA` which stands for Hawaiian Airlines. This is understandable as they 
have the longest leg to fly. 

The slowest carrier is `YV` which stands for Mesa Airlines. This airline does not seem to be a 
major airline and explains the leisurely pace at which the planes fly. 


```{r avg_flight_speed}
avg_flight_speed <- flights %>%
  group_by(carrier) %>%
  mutate(speed = mean(distance / air_time * 60, na.rm = TRUE)) %>%
  distinct(dest, .keep_all = TRUE) %>%
  select(carrier, speed)

filter(avg_flight_speed, speed == max(avg_flight_speed$speed))
filter(avg_flight_speed, speed == min(avg_flight_speed$speed))

filter(airlines, carrier == "HA")
filter(airlines, carrier == "YV")
```

### (h) (Bonus, not required) Find flights on which weekday (from Monday to Sunday) had the longest departure delay on average.

Done. Is there a better way to do this problem? I think I'm rather inefficient.

```{r longest_dep_weekday}
longest_dep_weekday <- flights %>%
  mutate(day_of_week = weekdays(time_hour)) %>%
  mutate(flight_number = paste(carrier, as.character(flight))) %>%
  group_by(day_of_week, flight_number) %>%
  summarize(day_of_week, flight_number,
            avg_delay = mean(dep_delay, na.rm = TRUE)) %>%
  distinct(day_of_week, flight_number, .keep_all = TRUE)

days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

results <- list()

for (day in days) {
  result <- longest_dep_weekday %>%
    filter(day_of_week == day) %>%
    arrange(desc(avg_delay)) %>%
    head(1)
  results[[day]] <- result
}

for (day in days) {
  cat("For", day, ":", "\n")
  print(results[[day]])
  cat("\n")
}

```

## 3. Analyzing the seattlepets data set


For the following questions, analyze the data set seattlepets in the package openintro. Read the help
document and make sure that you understand the basic information about the data set before analysis. 

### (a) How many species are there in the data set? What are they?

It looks like there are a lot of dogs and cats in seattle, but also goats and pigs..? 
There are 4 species in total.


```{r species_count}
species_count <- seattlepets %>%
  group_by(species) %>%
  summarize(count = n())

species_count
```

### (b) What are the three most common pet names in Seattle?


According to the data provided, Lucy, Charlie, and Luna are the most common pet names in Seattle.


```{r pet_names}
pet_names <- seattlepets %>%
  group_by(animal_name) %>%
  filter(!is.na(animal_name)) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

pet_names[1:3, ]
```

### (c) What are the ten most common pet names for cats? What are the ten most common pet names for dogs? Write a code to print the result and their frequencies.


Names like Lucy, Luna, Charlie, etc. are all really popular names. 

```{r dog_cat_common_names}
common_names <- seattlepets %>%
  filter(!is.na(animal_name)) %>%
  group_by(animal_name, species) %>%
  summarize(count = n())

species <- c("Dog", "Cat")

count_table <- list()

for (i in species) {
  temp <- common_names %>%
    filter(species == i) %>%
    arrange(desc(count)) %>%
    head(10)
  count_table[[i]] <- temp
}

for (i in species) {
  cat("For", i, ":", "\n")
  print(count_table[[i]])
  cat("\n")
}

count_table
```

### (d) How many names appear more than 100 times in the data set excluding “NA”?

There are 56 names that appear more than 100 times in the seattlepets dataset. 

```{r names_freq}
names_freq <- seattlepets %>%
  group_by(animal_name) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  filter(count >= 100 & animal_name != "NA")

nrow(names_freq)
```

### (e) For all names that appear more than 100 times in the data set, which has the highest “cat_to_dog” ratio? Which has the lowest? The “cat_to_dog” ratio can be computed this way - if a name appears 200 times, in which 150 are for cats and 50 are for dogs, the ratio is 150/50 = 3.

While Riley had the highest dog to cat ratio, Shadow had the highest cat to dog ratio. 

This means that Riley is way more likely to be used for dogs (understandable because riley is 
somewhat of a wild name and suitable for dogs) while shadow is more likely to be used for cats. 

```{r cat_to_dog_ratio}
#create a dataframe with the count given the same species and name
count_species_name <- seattlepets %>%
  filter(!is.na(animal_name)) %>%
  group_by(animal_name, species) %>%
  summarize(count = n())

#copy the code from above to see the 100 most common pet names in seattle
names_freq <- seattlepets %>%
  group_by(animal_name) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  filter(count >= 100 & animal_name != "NA")

#initialize a list with 2 subdataframes (dog/cat) that will count the frequency of names for both types of animals
sep_species <- list()

#initialize a new dataframe called ratios where results will be funneled in for top 100 pets
ratio <- data.frame(
  Name <- character(),
  cat_to_dog <- numeric(),
  dog_to_cat <- numeric()
)

#add entries to sep_species list
for (i in c("Dog", "Cat")) {

#for each species in the dataframe, filter out the dataframe and add it to the sep_species dataframe
  sep_species[[i]] <- count_species_name %>%
    filter(species == i)
}

#make a another for loop that adds a new entry to the ratio dataframe for all the most popular pet names
for (name in names_freq$animal_name) {

    count_dog <- sep_species$Dog %>% filter(animal_name == name)
    count_cat <- sep_species$Cat %>% filter(animal_name == name)

    new_data <- data.frame(
        Name = name,
        cat_to_dog = (count_cat$count / count_dog$count),
        dog_to_cat = (count_dog$count / count_cat$count)
    )
    ratio <- rbind(ratio, new_data)
}

#filter out all the names
ratio <- ratio %>%
  arrange(cat_to_dog)

head(ratio, 1)
tail(ratio, 1)
```
