---
title: "STA305 MIDTERM EXAM #1"
author: 'Jacky Guan'
date: 'Oct 14, 2023'
output:
  pdf_document: default
  html_notebook:
    fig_height: 4
    fig_width: 7
---



## Academic Honesty Statement (fill your name in the blank)

I, Mingjia (Jacky) Guan, hereby state that I have not communicated with or gained information in any way from my classmates or anyone other than the Professor during this exam, and that all work is my own.

## Load packages

```{r load-packages, message=FALSE}
library(tidyverse)
library(nycflights13)
library(openintro)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```


<br>

## 1. The `nycflights13` data set

### Question a)

<br>

The number of fligths departing from NYC each month roughly remains the same throughout the year. 


<br>

```{r hist_each_month}
dep_mon <- flights %>%
  group_by(month) %>%
  group_by(dep_time) %>%
  filter(!is.na(dep_time)) %>%
  summarise(dep_time, month) %>%
  mutate(dep_time, hour = dep_time %/% 100) %>%
  mutate(hour = ifelse(hour == 24, 0, hour))

ggplot(dep_mon, aes(x = month)) +
  geom_histogram(aes(fill = as.factor(hour)), binwidth = 0.5) +
  scale_x_continuous(breaks = 1:12) +
  scale_y_continuous(breaks = seq(0, 30000, by = 2000)) +
  labs(x = "Month of departure", y = "Number of flights",
  title = "Number of flights from NYC, by Month",
  fill = "Hour of departure") +
  theme(
  plot.title = element_text(hjust = 0.5, size = 20),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)
```

<br>

### Question b)

<br>

The top 10 destinations from NYC were 
- Chicago 
- Atlanta
- Los Angeles
- Boston
- Orlando
- Charlotte
- San Francisco
- Fort Lauderdale
- Miami
- Washington D.C.


<br>

```{r dest_10}
dest_10 <- flights %>%
  filter(!is.na(dest)) %>%
  group_by(dest) %>%
  summarize(dest, count = n()) %>%
  distinct(dest, .keep_all = TRUE) %>%
  arrange(desc(count)) %>%
  head(10)

airport_table <- data.frame()

for (dest in dest_10$dest) {
  airport_full_name <- airports %>%
    filter(faa == dest) %>%
    select(faa, name)

  airport_table <- bind_rows(airport_table, airport_full_name)
}

print(dest_10)
print(airport_table)
```

<br>

### Question c)

<br>

According to the graph and table, June has the worst delays. The average delay for June is 49.8 minutes.
July is a close second with an average delay of 48.8 minutes.

<br>

```{r dep_delays}
dep_delay <- flights %>%
  filter(dep_delay > 0) %>%
  group_by(month) %>%
  summarize(month, mean_delay = mean(dep_delay)) %>%
  distinct() %>%
  arrange(desc(mean_delay))

ggplot(dep_delay, aes(x = month, y = mean_delay)) +
  geom_bar(stat = "identity", aes(fill = as.factor(month))) +
  scale_x_continuous(breaks = 1:12) +
  scale_y_continuous(breaks = seq(0, 50, by = 5)) +
  labs(x = "Month of departure", y = "Average delay of flights (min)",
  title = "Average delay time of flights",
  fill = "Month") +
  theme(
  plot.title = element_text(hjust = 0.5, size = 20),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)

tibble(dep_delay)
```

<br>

### Question d)

<br>

The hottest days for the three aiports all occured in July. 
<br>
At JFK, the hottest day was on July 16th with a mean temperature of 87.4 degrees.
<br>
At EWR, the hottest day was on July 19th with a mean temperature of 91.1 degrees.
<br>
At LGA, the hottest day was also on July 19th with a mean temperature of 92.1 degrees.

<br>

```{r mean_temp}
mean_temp <- weather %>%
  group_by(month, day, origin) %>%
  summarize(month, day, origin, mean_temp = mean(temp)) %>%
  distinct() %>%
  arrange(desc(mean_temp)) %>% 
  head(9) %>%
  filter(day == 19 | day == 16)

tibble(mean_temp)
```

<br>

### Question e)

<br>

Overnight flights usually fly an average of 201 minutes (3 hours and 21 minutes).
These flights are longer than non-overnight flights, which have an average of 152 minutes. 

<br>

```{r overnight}
overnight <- flights %>%
  filter(dep_time > 2100 & between(arr_time, 300, 700))

non_overnight <- flights %>%
  filter(dep_time <= 2100 & !between(arr_time, 300, 700))

mean(overnight$air_time, na.rm = TRUE)
mean(non_overnight$air_time, na.rm = TRUE)
```

<br>

### Question f)

<br>

The temperature seems to increase as average delay increases. At the same time, visibility does not 
seem to be a significant factor in terms of average delays. 

<br>

BONUS
```{r bonus}
dep_delay_day <- flights %>%
  filter(dep_delay > 0) %>%
  group_by(month, day) %>%
  summarize(month, day, mean_delay = mean(dep_delay)) %>%
  distinct()

mean_temp_simple <- weather %>%
  group_by(month, day) %>%
  summarize(month, day, mean_temp = mean(temp)) %>%
  distinct()

mean_visibility <- weather %>%
  group_by(month, day) %>%
  summarize(month, day, mean_visibility = floor(mean(visib))) %>%
  distinct()

combined_data <- left_join(dep_delay_day, mean_temp_simple,
  by = c("month", "day"))
combined_data <- left_join(combined_data, mean_visibility,
  by = c("month", "day"))

ggplot(combined_data, aes(mean_delay, mean_temp)) +
  geom_point(aes(color = as.factor(mean_visibility))) +
  geom_smooth() +
  labs(x = "Average Delay (min)", y = "Average Temperature (F)",
  title = "The effect of temperature on delay times",
  labs = "Visibility") +
  scale_y_continuous(breaks = seq(20, 100, by = 5)) +
  scale_x_continuous(breaks = seq(0, 100, by = 5)) +
  theme(
  plot.title = element_text(hjust = 0.5, size = 20),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)
```

<br>

<br>

## 2. The `babies` data set

### Question a)

<br>

The distribution of gestation based on parity seems to be that of a normal distribution.
There is no significant difference between gestation based on parity.

<br>


```{r parity_gestation}
parity_gestation <- babies %>%
  mutate(parity, first_pregnancy_flag =
           ifelse(parity == 0, "First Pegnancy", "Not First Pregnancy"))

ggplot(parity_gestation, aes(x = first_pregnancy_flag, y = gestation)) +
  geom_boxplot(aes(fill = first_pregnancy_flag)) +
  labs(x = "Parity", y = "Gestation Period (days)",
  title = "Gestation period by Parity",
  fill = "Parity") +
  scale_y_continuous(breaks = seq(200, 340, by = 10), limits = c(200, 340)) +
  theme(
  plot.title = element_text(hjust = 0.5, size = 20),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)
```

<br>

### Question b)

<br>

There does not seem to be a significant visual 
difference between gestation period and whether the mother was a smoker or not.

<br>

```{r smoke}
smoke <- babies %>%
  mutate(smoker_flag = case_when(
    smoke == 0 ~ "Smoker",
    smoke == 1 ~ "Non-Smoker",
    is.na(smoke) ~ "Unknown"
    )
  )

ggplot(smoke, aes(smoker_flag, gestation)) +
  geom_boxplot(aes(fill = smoker_flag)) +
  labs(x = "Smoking Status", y = "Gestation period (days)",
  title = "The effect of smoking on gestation period",
  fill = "Smoking Status") +
  scale_y_continuous(breaks = seq(225, 325, by = 25), limits = c(225, 325)) +
  theme(
  plot.title = element_text(hjust = 0.5, size = 20),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)
```

<br>

### Question c)

<br>

Age does not seem to be a factor when considering gestation period as the line of best fit is flat.

<br>

```{r gestation_age}
ggplot(babies, aes(age, gestation)) +
  geom_point(position = "jitter", aes(color = as.factor(gestation))) +
  geom_smooth() +
  guides(color = FALSE) +
  labs(x = "Age", y = "Gestation period (days)",
  title = "The effect of age on gestation period",
  fill = "Legend Title") +
  scale_y_continuous(breaks = seq(225, 350, by = 25), limits = c(225, 350)) +
  scale_x_continuous(breaks = seq(15, 45, by = 3)) +
  theme(
  plot.title = element_text(hjust = 0.5, size = 20),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)
```

<br>

### Question d)

<br>

While parity does not seem to have a significant effect on birthweight, gestation period does. 
The longer the gestation period, the heavier the child tends to be. 

<br>

```{r gestation_pregnancy}
gestation_pregnancy <- babies %>%
  mutate(parity, first_pregnancy_flag =
           ifelse(parity == 0, "First Pegnancy", "Not First Pregnancy")) %>%
  filter(!is.na(gestation)) %>%
  mutate(gestation, gest_category = case_when(
    between(gestation, 0, 250)  ~ "<250",
    between(gestation, 250, 270)  ~ "250-270",
    between(gestation, 270, 290)  ~ "270-290",
    between(gestation, 290, 310)  ~ "290-310",
    between(gestation, 310, 330)  ~ "330-350",
    between(gestation, 330, 1000)  ~ ">350"
    )
  )

ggplot(gestation_pregnancy, aes(gest_category, bwt)) +
  geom_boxplot(aes(fill = first_pregnancy_flag)) +
  labs(x = "Gestation period (days)", y = "Birthweight (ounces)",
  title = "Effect of gestation and parity on Birthweight",
  fill = "Parity") +
  theme(
  plot.title = element_text(hjust = 0.5, size = 20),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)
```

<br>

### Question e)

<br>

While the BMI of the mother didn't seem to have a significant impact on birthweight, 
it seems that if the mother smoked, the baby would likely weigh more.

<br>

```{r smoker_bmi}
smoker_bmi <- babies %>%
  mutate(smoker_flag = case_when(
    smoke == 0 ~ "Smoker",
    smoke == 1 ~ "Non-Smoker",
    is.na(smoke) ~ "Unknown"
    )
  ) %>%
  mutate(weight, height, bmi = (weight / (height**2)) * 703) %>%
  mutate(bmi, bmi_category = case_when(
    between(bmi, 0, 18.5)  ~ "Under 18.5",
    between(bmi, 18.5, 24.9)  ~ "18.5-24.9",
    between(bmi, 25, 29.9)  ~ "25-29.9",
    between(bmi, 30, 39.9)  ~ "30-39.9",
    bmi >= 40  ~ "40 or over",
    TRUE ~ "Unknown"
  ))

ggplot(smoker_bmi, aes(bmi_category, bwt)) +
  geom_boxplot(aes(fill = smoker_flag)) +
  labs(x = "BMI", y = "Birthweight (ounces)",
  title = "Effect of Smoking and Mother's BMI on birthweight",
  fill = "Smoking Status") +
  theme(
  plot.title = element_text(hjust = 0.5, size = 20),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)
```

<br>
