---
title: "STA 305 HW6"
author: "Jacky Guan"
date: "2023-12-05"
output: pdf_document
---


Loading Required libraries
```{r setup, message=FALSE}
library(tidyverse)
library(openintro)
library(nycflights13)
library(lubridate)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```


## 1. Date and Time - nycflights13 data set




#### How many timezones are there for all the destination airports (excluding NA)?


There are a total of 6 rows when it comes to destination airports.

```{r timezones}
timezones <- flights %>%
  select(dest) %>%
  filter(!is.na(dest)) %>%
  left_join(airports, c("dest" = "faa")) %>%
  filter(!is.na(tz)) %>%
  select(tz) %>%
  group_by(tz) %>%
  mutate(count = n()) %>%
  distinct(tz, .keep_all = TRUE)

print(timezones)
print(nrow(timezones))
```





#### Use the data in flights and airports only along with R code to show the time difference (in hours) between New York City and the following cities:


* Chicago

Chicago is 1 hour behind NY.

```{r chicago_ny}
chicago_ny <- flights %>%
  select(dest, origin) %>%
  filter(dest == "ORD") %>%
  left_join(airports, c("dest" = "faa")) %>%
  left_join(airports, c("origin" = "faa")) %>%
  distinct(dest, .keep_all = TRUE)

delta_t <- chicago_ny$tz.x - chicago_ny$tz.y
print(delta_t)
```


* Dallas

Dallas is 1 hour behind NY.

```{r dallas_ny}
dallas_ny <- airports %>%
  filter(str_detect(name, "Newark|Dallas")) %>%
  distinct(tz, .keep_all = TRUE) %>%
  mutate(delta_t = tz - lag(tz)) %>%
  filter(!is.na(delta_t))

print(dallas_ny$delta_t)
```

* Denver

Denver is 2 hours behind NY.

```{r denver_ny}
denver_ny <- airports %>%
  filter(str_detect(name, "Denver|Newark")) %>%
  distinct(tz, .keep_all = TRUE) %>%
  mutate(delta_t = tz - lag(tz)) %>%
  filter(!is.na(delta_t))

print(denver_ny$delta_t)
```

* Seattle

Seattle is 3 hours behind NY.

```{r seattle_ny}
seattle_ny <- airports %>%
  filter(str_detect(name, "Seattle|Newark")) %>%
  distinct(tz, .keep_all = TRUE) %>%
  mutate(delta_t = tz - lag(tz)) %>%
  filter(!is.na(delta_t))

print(seattle_ny$delta_t)
```
* Anchorage 

Anchorage is 4 hours behind NY.


```{r anchorage_ny}
anchorage_ny <- airports %>%
  filter(str_detect(name, "Anchorage|Newark")) %>%
  distinct(tz, .keep_all = TRUE) %>%
  mutate(delta_t = tz - lag(tz)) %>%
  filter(!is.na(delta_t))

print(anchorage_ny$delta_t)
```

* Honolulu

Honolulu is 5 hours behind NY.

```{r honolulu_ny}
honolulu_ny <- airports %>%
  filter(str_detect(name, "Honolulu|Newark")) %>%
  distinct(tz, .keep_all = TRUE) %>%
  mutate(delta_t = tz - lag(tz)) %>%
  filter(!is.na(delta_t))

print(honolulu_ny$delta_t)
```

#### Write a function Time_difference_NYC(dest) to compute the time difference (in hours) between any destination airport (in faa code) and New York City.

```{r Time_difference_NYC}
Time_difference_NYC <- function(dest) {
  return <- airports %>%
    filter( faa == dest | faa == "JFK") %>%
    distinct(tz, .keep_all = TRUE) %>%
    mutate(delta_t = tz - lag(tz)) %>%
    filter(!is.na(delta_t))
  return$delta_t
}

Time_difference_NYC("0S9") # Jefferson County International Airport
Time_difference_NYC("JFK") # JFK in NYC
Time_difference_NYC("SLC") # Salt Lake City
Time_difference_NYC("AUS") # Austin Texas
Time_difference_NYC("LAX") # Los Angeleles
```


### Write a function flight_time(dep_time, arr_time, origin, dest) to compute the actual flight time from the dep_time at the origin airport to the arr_time in the dest airport. 
dep_time and arr_time should be in the form of an integer such as 830 (8:30am); origin and dest should be in faa codes.

```{r travel_time}
flight_time <- function(dep_time, arr_time, origin, dest) {
  raw_flight_time <- arr_time - dep_time

  origin_time <- airports %>%
    filter(faa == origin) %>%
    select(tz)

  dest_time <- airports %>%
    filter(faa == dest) %>%
    select(tz)

  delta_t <- dest_time - origin_time
  numeric_time <- raw_flight_time + delta_t * 100

  #for the case when hours and 
  if (numeric_time %% 100 >= 60) {
    numeric_time <- numeric_time + 40
  }

  hour <- as.character(numeric_time %/% 100)
  minute <- as.character(numeric_time %% 100)

  x <- paste(hour, minute)

  time_hm <- hm(x)

  formatted_time <- format(time_hm, format = "%I:%M%p")

  return(formatted_time)
}

flight_time(1234, 2334, "JFK", "LAX")
flight_time(815, 1545, "ORD", "SFO")
flight_time(945, 1745, "ATL", "MIA")
flight_time(1215, 2045, "DFW", "SEA")
flight_time(1325, 2155, "DEN", "MSP")
flight_time(1030, 1845, "LGA", "SJC")
```


## 2. Lists and Loops
#### Write a code to create a list that contains 100 vectors, where the first vector is 1 alone, the second is 1 and 2, the third is 1,2,3 etc.

```{r looping}
my_list <- lapply(1:100, function(n) seq(1:n))

print(head(my_list))
```


## 3. Functional
Use map function to fulfill the following tasks from your midterm project data set. 

#### First, create a numeric column BMI as instructed from the midterm project. Also create a categorical column obesity_flag which has the value of normal or underweight or overweight or obese depending on the categories for NObeyesdad.

```{r obesity}
setwd("/Users/jacky/Documents/SEM 3/STA305/Homework/HW6")
obesity <- read.csv("obesity_dataset.csv")
```

```{r obesity_reformatting}
obesity <- obesity %>%
  mutate(Weight, Height, bmi = (Weight / (Height**2)))

obesity <- obesity %>%
  mutate(NObeyesdad = fct_reorder(NObeyesdad, bmi)) %>%
  rename(obesity_level = NObeyesdad) %>%
  rename(
    gender = Gender,
    age = Age,
    height = Height,
    weight = Weight,
    fam_hist = family_history_with_overweight,
    high_cal = FAVC,
    freq_vegetable = FCVC,
    meal_count = NCP,
    snack = CAEC,
    activity = FAF,
    smoke = SMOKE,
    water = CH2O,
    monitor_calorie = SCC,
    screen_time = TUE,
    alcohol = CALC,
    transportation = MTRANS
  ) %>%
  mutate(
    obesity_flag = case_when(
      obesity_level %in% c("Insufficient_Weight",
                           "Normal_Weight") ~ "Normal or Underweight",
      obesity_level %in% c("Overweight_Level_I",
                           "Overweight_Level_II") ~ "Overweight",
      obesity_level %in% c("Obesity_Type_I",
                           "Obesity_Type_II", "Obesity_Type_III") ~ "Obese",
      TRUE ~ NA_character_
    )
  )
```



#### • For all numeric variables (other than BMI and weight), compute the correlation coefficient with BMI, and sort your variables by your result in a descending order. Which variable has the biggest correlation coefficient? 
Also explain the result for height. Why is the correlation coefficient a small number although BMI is directly calculated from height (and weight)?






I believe that height doesn't play such a large role as height won't really determine one's weight, which has a 
larger impact on BMi. As the name implies, BMI stands for body mass index and not body height index.


```{r numeric_testing}
numeric_vars <- sapply(obesity, is.numeric)
numeric_obesity_vars <- obesity[, numeric_vars]

bmi <- obesity$bmi

cor_results <- lapply(colnames(numeric_obesity_vars), function(col) {
  correlation_result <- cor(obesity[[col]], bmi)
  tibble(Variable = col, cor_coeff = correlation_result)
})

cor_results_df <- bind_rows(cor_results)

cor_results_df <- cor_results_df %>%
  arrange(desc(cor_coeff))

print(cor_results_df)
```


#### • For all categorical variables (other than the target variable), compute the p-value when doing the chi- square test between each of them and obesity_flag, and sort your variables by the result in an ascending order. Which variable is most relevant (smallest p-value)?



It seems that snacking is the most relevant variable when it comes to correlation with BMI.


```{r categorical_testing}
cat_vars <- sapply(obesity, is.character)
cat_obesity_vars <- obesity[, cat_vars]
obesity_flag <- obesity$obesity_flag

cat_results <- list()

for (col in colnames(cat_obesity_vars)) {
  contingency_table <- table(obesity[[col]], obesity_flag)
  correlation_result <- chisq.test(contingency_table)
  cat_results[[col]] <- correlation_result
}

cat_results_df <- bind_rows(lapply(cat_results, broom::tidy), .id = "Variable")


cat_results_df <- cat_results_df %>%
  arrange(p.value) %>%
  print()
```


#### • For all numeric variables (other than BMI, weight and height), compute the p-value when doing the t-test between each of them and obesity_flag, and sort your in a descending order.


```{r t_testing}
num_vars <- sapply(obesity, is.numeric)
num_obesity_vars <- obesity[, num_vars]
obesity_flag <- obesity$obesity_flag

num_results <- list()

for (col in colnames(num_obesity_vars)) {
  contingency_table <- table(obesity[[col]], obesity_flag)
  correlation_result <- t.test(contingency_table)
  num_results[[col]] <- correlation_result
}
num_results_df <- bind_rows(lapply(num_results, broom::tidy), .id = "Variable")


num_results_df <- num_results_df %>%
  arrange(desc(p.value)) %>%
  print()
```