---
title: "Diamonds - Class Notes"
author: "Jacky Guan"
date: "2023-09-19"
output: html_document
---

#### Lab Exercise for `diamonds` data set
```{r setup, include=FALSE}
library(tidyverse)
```

### How many samples are there? How many variables are there?

Samples and Variables?

```{r samples/variables}
ncol(diamonds)
nrow(diamonds)
```

### Understand the meaning and the data type for each variable.

The number of colums in diamonds is how many attributes diamonds have. The number of rows is how many diamonds exist in the data set.

### Create a plot to study the relationship between `price` and `carat`. How do you understand this plot?

```{r Carat}
ggplot(diamonds) +
    geom_point(aes(x = carat, y = price))
```

It seems that the price of diamonds grows exponentially with an increase in carats. 
The plot shows trend lines along carat%(0.5), indicating that most people wish to purchase diamonds with memorable carat numbers

### Create plots to study the relationship between `price` and `cut`, `color`, `clarity`, respectively. How do you understand these plots?

It seems that these plots individually will not provide the whole picture.

#### `cut`
```{r Cut}
ggplot(diamonds, aes(x = cut, y = price)) +
    geom_boxplot(aes(fill = cut))
```

#### `color`
```{r Color}
ggplot(diamonds, aes(x = color, y = price)) +
    geom_boxplot(aes(fill = color))
```

#### `clarity`
```{r Clarity}
ggplot(diamonds, aes(x = clarity, y = price)) +
    geom_boxplot(aes(fill = clarity))
```

### After finishing 3 and 4, do you have more questions raised in your mind? Create a plot to answer those questions.

I do have some questions remaining. What is the relationship betweeen three variables at a time?

#### Carat, Cut, and Price

```{r Carat_Cut}
ggplot(diamonds, aes(x = carat)) +
    geom_bar(aes(fill = cut))
```

#### Carat, Color, and Price

```{r Carat_Color}
ggplot(diamonds, aes(x = carat)) +
    geom_bar(aes(fill = color))
```

#### Carat, Clarity and Price

```{r Carat_Clarity}
ggplot(diamonds, aes(x = carat)) +
    geom_bar(aes(fill = clarity))
```

#### Cut, Color and Price

```{r Cut_Color}
ggplot(diamonds, aes(x = cut, y = price)) +
    geom_boxplot(aes(fill = color))
```

#### Cut, Clarity and Price

```{r Cut_Clarity}
ggplot(diamonds, aes(x = cut, y = price)) +
    geom_boxplot(aes(fill = clarity))
```
#### Color, Clarity and Price

```{r Color_Clarity}
ggplot(diamonds, aes(x = color, y = price)) +
    geom_boxplot(aes(fill = clarity))
```

For this specific data set, the IF clarity in the D color range is a large outlier. 

Let's analyze that specific category. TBC.
```{r IF/D}
filter(diamonds, diamonds$clarity == "IF" , diamonds$color == "D")
```

