---
title: "Logical Operators and Flight Data - Class Notes"
author: "Jacky Guan"
date: "2023-09-21"
output: html_document
---

#### Lab Exercise for `nycflights13` data set
```{r setup}
library(tidyverse)
library(nycflights13)
tibble(flights)
```

## `near()` to approximate for boolean operators

```{r near}
sqrt(2)^2 == 2
near(sqrt(2)^2, 2)
```

## Comparing with `&` and `&&``

While `&` compares data entries by index, `&&` will only compare data sets as a whole.

```{r &}
c(5, 2, 3) & c(5, 1, 1)
c(5, 2, 3) && c(5, 1, 1)

```

The warning message forces non-zero vectors into a `True` statement. 

# Lab Exercises

## a) Find all flights that depart from JFK and land on ORD or CVG (Chicago or Dallas).

```{r JFK_D/C}
l1 <- filter(flights, origin == "JFK", dest == "ORD" | dest == "CVG")
tibble(l1)
```

## b) Find all flights that flied a distance between 200 and 500 miles.

*Note that one cannot use the between operator like in python. One has to use the `between` operator*

```{r 2/5}
l2 <- filter(flights, between(distance, 200, 500))
tibble(l2)
```

## c) Plot the histogram of departure time for filtered data either in a) or b).

### For a)

```{r hist.a}
ggplot(l1) +
    geom_histogram(aes(x = dep_time/100, fill = carrier), binwidth = 1)
```

## How many flights have a missing plane tail number? What is the percentage of flights with a missing plane tail number?

```{r missing_tail_number}
tailna <- filter(flights, is.na(tailnum))
nrow(tailna)/nrow(flights)
```


## For `EV`, during which month were most flights canceled?

```{r canceled_flights}
canceled_flights <- filter(flights, is.na(dep_time))
ggplot(canceled_flights) + 
  geom_bar(aes(carrier)) +
  labs(title = "Canceled Flights by Carrier", 
       x = "Carrier (in abbreviation)",
       y = "Counts")


ggplot(canceled_flights) +
  geom_bar(aes(month))
```

## What was the longest travel distance for any flight in our data set? What was the origin and the destination?

```{r longest_flight}
arrange(flights, desc(air_time))
```

The longest flight was from `EWR` to `HNL`. Where is `HNL`? Intuitionally, it should be Hawaii.

```{r hnl}
filter(airports, faa == "HNL")
```

`HNL` is going to be Honolulu International Airport. 

## What was the shortest air time for any flight (that actually finished the trip) in our data set? What was the origin and the destination?

```{r short_airtime}
arrange(flights, desc(air_time))
```

The `origin` was from `EWR` to `BDL`. Where is `BDL`?

```{r bdl}
filter(airports, faa == "BDL")
```

`BDL` is going to be Bradley International Airport. 